---
title: "Practical_Machine_learning"
author: "JC"
date: "December 18, 2017"
output: html_document
---

#Practical Machine Learning Course Project Report

These is a file produced during a homework assignment of Coursera's MOOC Practical Machine Learning from Johns Hopkins Bloomberg School of Public Health.
For more information about the several MOOCs comprised in this Specialization, please visit: https://www.coursera.org/specialization/jhudatascience/



GitHub Repo: https://github.com/JamesCanterbury/Practical_Machine_Learning_Course_Project

Rpub: http://rpubs.com/

#Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Data set).

#Data Sources

The training data for this project is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project comes from this original source: 

http://groupware.les.inf.puc-rio.br/har. 

If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.
Intended Results

# Intended Results 
The goal of this project is to predict the manner in which they did the exercise. This is the "class" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HMTL file describing your analysis. Please constrain text to less than 2000 words and the number of figures to be less than 5. It will make it easier for graders. 

2. You should also apply your machine learning algorithm to 20 test cases available in the test data above. Please submit your predictions in the appropriate format. See the programming assignment for additional details. 

#Reproducibility

In order to reproduce the same results, you need a certain set of packages as well as setting a pseudo random seed equal to the one I have used.
Note: To install, for instance, the rattle  package in R, run this command:  install.packages("rattle").
The following Libraries were used for this project, which you should install and load them in your working environment.

```{r setup, include=FALSE, message=FALSE}


 setwd("/home/mpiuser/Desktop/ml_coursera")

knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE,}

library(rattle)
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(randomForest)
library(RColorBrewer)
library(e1071  )
```

Finally, load the same seed with the following line of code:

```{r}
set.seed(12345)

```

## Getting Data

The following code fragment downloads the data set to the data folder in the current working directory.

```{r download data}

trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "pml-training.csv"
testFile  <- "pml-testing.csv"

if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile = trainFile, method = "wget")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile = testFile, method = "wget")
}

rm(trainUrl)
rm(testUrl)

training <- read.csv(trainFile)
test <- read.csv(testFile)


```

## Reading Data

After downloading the data from the data source, we can read the two csv files into two data frames.

```{r reading data, echo=TRUE}
training <- read.csv(trainFile)
test <- read.csv(testFile)

dim(training)
dim(test)
rm(testFile, trainFile)

```

The training data set contains `r dim(training)[1]` observations and `r dim(training)[2]`  variables, while the testing data set contains `r dim(test)[1]` observations and `r dim(test)[2]` variables. The class variable in the training set is the outcome to predict.

#Cleaning Data

In this step, we will clean the data set and get rid of observations with missing values as well as some meaningless variables.

###1. We clean the Near Zero Variance Variables.

```{r cleaning data}
NZV <- nearZeroVar(training, saveMetrics = TRUE)
head(NZV, 20)
training01 <- training[, !NZV$nzv]
testing01 <- test[, !NZV$nzv]
dim(training01)
dim(testing01)
rm(training)
rm(test)
rm(NZV)

```

###2.     Removing some columns of the dataset that do not contribute much to the accelerometer measurements.

```{r warning= FALSE}
regex <- grepl("^X|timestamp|user_name", names(training01))
training <- training01[, !regex]
testing <- testing01[, !regex]
rm(regex)
rm(training01)
rm(testing01)
dim(training)
dim(testing)
```
    
###3. Removing columns that contain NA's.

```{r}
cond <- (colSums(is.na(training)) == 0)
training <- training[, cond]
testing <- testing[, cond]
rm(cond)
```

Now, the cleaned training data set contains `r dim(training)[1]`  observations and `r dim(training)[2] ` variables, while the testing data set contains ` r dim(testing)[1] ` observations and `r dim(testing)[2] ` variables.

Correlation Matrix of Columns in the Training Data set.

```{r}
corrplot(cor(training[, -length(names(training))]), method = "color", tl.cex = 0.5)
```

#Partitioning Training Set

we split the cleaned training set into a pure training data set (70%) and a validation data set (30%). We will use the validation data set to conduct cross validation in future steps.

```{r}
set.seed(56789) # For reproducibile purpose
inTrain <- createDataPartition(training$classe, p = 0.70, list = FALSE)
validation <- training[-inTrain, ]
training <- training[inTrain, ]
rm(inTrain)
```

The Data set now consists of `r dim(training)[2]` variables with the observations divided as following:

    Training Data: `r dim(training)[1]` observations.
    Validation Data: `r dim(validation)[1]` observations.
    Testing Data: `r dim(testing)[1]` observations.

#Data Modelling

##Decision Tree

We fit a predictive model for activity recognition using Decision Tree algorithm.

```{r modeltree}
modelTree <- rpart(classe ~ ., data = training, method = "class")
prp(modelTree)
```

Now, we estimate the performance of the model on the validation data set.

```{r tree}
predictTree <- predict(modelTree, validation, type = "class")
confusionMatrix(validation$classe, predictTree)
accuracy <- postResample(predictTree, validation$classe)
acc <- accuracy *100
acc <- format(round(acc, 2))
ose <- 1 - as.numeric(confusionMatrix(validation$classe, predictTree)$overall[1])
ose <- format(round((ose * 100),2))
rm(predictTree)
rm(modelTree)
```

The Estimated Accuracy of the Random Forest Model is `r acc[1]`%  and the Estimated Out-of-Sample Error is `r ose`% .

#Random Forest

We fit a predictive model for activity recognition using Random Forest algorithm because it automatically selects important variables and is robust to correlated covariates & outliers in general.
We will use 5-fold cross validation when applying the algorithm.

```{r randomforest run}
modelRF <- train(classe ~ ., data = training, method = "rf", trControl = trainControl(method = "cv", 5), ntree = 250) 
modelRF
```

Now, we estimate the performance of the model on the validation data set.

```{r}
predictRF <- predict(modelRF, validation)
confusionMatrix(validation$classe, predictRF)
accuracy <- postResample(predictRF, validation$classe)
ose <- 1 - as.numeric(confusionMatrix(validation$classe, predictRF)$overall[1])
rm(predictRF)
```

The Estimated Accuracy of the Random Forest Model is `r accuracy[1]` *100% and the Estimated Out-of-Sample Error is `r ose` *100%`.
Random Forests yielded better Results, as expected!

Predicting The Manner of Exercise for Test Data Set

Now, we apply the Random Forest model to the original testing data set downloaded from the data source. We remove the problem_id column first.

```{r}
rm(accuracy)
rm(ose)
predict(modelRF, testing[, -length(names(testing))])
```

#Generating Files to submit as answers for the Assignment

Function to generate files with predictions to submit for assignment.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("./Assignment_Solutions/problem_id_",i,".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}

```

Generating the Files.

```{r}
pml_write_files(predict(modelRF, testing[, -length(names(testing))]))
rm(modelRF)
rm(training)
rm(testing)
rm(validation)
rm(pml_write_files)
```


